Reproducible Research, Peer Assessment 1
========================================================
## Importing & Processing Data
```{r import}
# Import data
activity <- read.csv("activity.csv")
# Aggregate by date
dailySteps <- aggregate(activity$steps, list(as.Date(activity$date)), sum)
# Name usefully
names(dailySteps) <- c("Date", "Total.Steps")
# Converts to numeric
dailySteps$Total.Steps <- as.numeric(dailySteps$Total.Steps)
```

## What is the mean total number of steps taken per day?
1. The daily totals are distributed like so:

    ```{r fig.width=7, fig.height=6}
    library(ggplot2)
    ggplot(dailySteps, aes(x=Total.Steps)) + geom_histogram(binwidth=1000, colour="black", fill="white")
    ```

    ```{r central}
    # Calculate mean
    dailyMean <- mean(dailySteps$Total.Steps, na.rm = TRUE)
    #Calculate median
    dailyMedian <- median(dailySteps$Total.Steps, na.rm = TRUE)
    ```

2. The mean daily number of steps is `r dailyMean`; the median daily number of steps is `r dailyMedian`.

## What is the average daily activity pattern?
    
1. The mean number of steps across all days, by five-minute interval:
    ```{r dailyCalc}
    intervalPattern <- aggregate(activity$steps, list(activity$interval), mean, na.rm = TRUE)
    names(intervalPattern) <- c("5min.Interval", "Average.Num.Steps")
    ```

    ```{r fig.width=7, fig.height=6}
    plot(intervalPattern, type="l", ylab = "Mean Number of Steps", xlab = "5min Interval")
    ```
2. The 5-minute interval with the greatest average number of steps is `r intervalPattern[intervalPattern$Average.Num.Steps == max(intervalPattern$Average.Num.Steps),][[1]]`.

## Missing Values
1. The total number of missing values in this dataset is `r sum(is.na(activity))`.

2. Missing values will be replaced with the overall mean.

3. This code replaces NAs with the overall mean.

    ```{r replaceMissing}
    for (i in which(sapply(activity, is.numeric))) {
        activity[is.na(activity[, i]), i] <- mean(activity[, i],  na.rm = TRUE)
    }
    ```

4. The new daily totals are distributed like so:

    ```{r dailyPostCalc}
    # Aggregate by date
    dailySteps <- aggregate(activity$steps, list(as.Date(activity$date)), sum)
    # Name usefully
    names(dailySteps) <- c("Date", "Total.Steps")
    # Converts to numeric
    dailySteps$Total.Steps <- as.numeric(dailySteps$Total.Steps)
    ```

    ```{r fig.width=7, fig.height=6}
    ggplot(dailySteps, aes(x=Total.Steps)) + geom_histogram(binwidth=1000, colour="black", fill="white")
    ```

    ```{r centralAdj}
    # Calculate mean
    dailyAdjMean <- mean(dailySteps$Total.Steps, na.rm = TRUE)
    #Calculate median
    dailyAdjMedian <- median(dailySteps$Total.Steps, na.rm = TRUE)
    ```

The new mean daily number of steps is `r dailyAdjMean`, compared to the original `r dailyMean`; the new median daily number of steps is `r dailyAdjMedian`, compared to the original `r dailyMedian`.  Replacing missing values with the overall mean distorts the distribution of the data, but since the original mean/median were calculated leaving out the NAs, does not materially change the estimate of central tendency.

## Are there differences in activity patterns between weekdays and weekends?
1. The factor daytype has been added to the data as below.

    ```{r daytype}
    activity$daytype <- factor(ifelse(as.POSIXlt(activity$date)$wday %% 6 == 0, "Weekend", "Weekday"))
    ```

2. The time series panel plot of average number of steps taken per 5 min interval per day type

    ```{r dayTimeSeries, warning=FALSE}
    library(reshape2)
    test <- melt(activity, c("interval", "daytype"))
    test <- test[1:17568,]
    test$value <- as.numeric(test$value)
    dayIntervalPattern <- dcast(test, interval ~ daytype, mean)
    ```

    ```{r fig.width=7, fig.height=6}
    library(lattice)
    px1 = xyplot(Weekday ~ interval, dayIntervalPattern, type = "l", ylab = "Number of Steps", main = "Weekday")
    px2 = xyplot(Weekend ~ interval, dayIntervalPattern, type = "l", ylab = "Number of Steps", main = "Weekend")

    # arrange the two plots vertically
    print(px1, position=c(0, .5, 1, 1), more=TRUE)
    print(px2, position=c(0, 0, 1, .5))
